---
title: "HW8"
author: "Brendan Graham"
date: "March 19, 2016"
output: word_document
---
1)

Using the classical decomposition technique from class, I fit an ARMA(3,2) model to the residuals of my regression. This gives me 


Using an ARMA(3,2) model with seasonal differencing results in the following forecast for the next year

```{r, echo = FALSE}
pred2=sarima.for(log(unemp), 12, 3, 1, 2, 0, 1, 0, 12)
```




4.1)

```{r, echo = FALSE, include = FALSE}

load("~/R Working Directory/Villanova/time_series_and_forecasting/project/phl/data/phl.rda")

```

a) Changing n from 100 to 128 reduces the frequencies in x1, x2 and x3 from 0.06, 0.10, 0.40 to 0.046875, 0.078125 and 0.3125, respectively. The increase in n doesn't affect how the plots of x1, x2, and x3 look, but it does change the plot of their sum.

```{r, echo=FALSE}

#4.1
#old
x1 = 2*cos(2*pi*1:100*6/100) + 3*sin(2*pi*1:100*6/100)
x2 = 4*cos(2*pi*1:100*10/100) + 5*sin(2*pi*1:100*10/100)
x3 = 6*cos(2*pi*1:100*40/100) + 7*sin(2*pi*1:100*40/100)
x.1 = x1 + x2 + x3
par(mfrow=c(2,2))
plot.ts(x1, ylim=c(-10,10), main=expression(omega==6/100~~~A^2==13))
plot.ts(x2, ylim=c(-10,10), main=expression(omega==10/100~~~A^2==41))
plot.ts(x3, ylim=c(-10,10), main=expression(omega==40/100~~~A^2==85))
plot.ts(x.1, ylim=c(-20,20), main="Original Sum")


#new
n <- 128
x1 = 2*cos(2*pi*1:n*6/n) + 3*sin(2*pi*1:n*6/n)
x2 = 4*cos(2*pi*1:n*10/n) + 5*sin(2*pi*1:n*10/n)
x3 = 6*cos(2*pi*1:n*40/n) + 7*sin(2*pi*1:n*40/n)
x.2 = x1 + x2 + x3
par(mfrow=c(2,2))
plot.ts(x1, ylim=c(-10,10), main=expression(omega==6/128~~~A^2==13))
plot.ts(x2, ylim=c(-10,10), main=expression(omega==10/128~~~A^2==41))
plot.ts(x3, ylim=c(-10,10), main=expression(omega==40/128~~~A^2==85))
plot.ts(x.2, ylim=c(-20,20), main="New Sum")

```


b) The change in frequency is present in the scaled periodograms. The old periodogram on top shows spikes corresponding to the original frequencies, 0.06, 0.10 and 0.40, while the new periodogram shows spikes at the new frequencies, 0.046875, 0.078125 and 0.3125. Overall, increasing N reduced the frequencies which is reflected in the 3 spikes shifted to lower frequencies in the new periodogam relative to the old periodogram.

```{r, echo = FALSE, fig.height=12, fig.width=12}
#old
P.1 = Mod(2*fft(x.1)/100)^2
Fr.1 = 0:99/100

#new
P.2 = Mod(2*fft(x.2)/n)^2
Fr.2 = 0:127/128

par(mfrow=c(2,1))
plot(Fr.1, P.1, type="o", xlab="frequency", ylab="Old periodogram")
plot(Fr.2, P.2, type="o", xlab="frequency", ylab="New periodogram")

```


c) The addition of the white noise terms...The addition of the white noise term also introduces some minor fluctuations across the periodogram at every frequency, but the 3 spikes at the specificed freqiencies are still present. The previous periodogram has no fluctuations ther than the spikes at the 3 specified frequencies.

```{r, echo = FALSE, warning = FALSE, fig.height=12, fig.width=12}

#part c
#old
x.1 = x1 + x2 + x3
plot.ts(x.1, ylim=c(-20,20), main="Original Sum")

#new
x1 = 2*cos(2*pi*1:100*6/100) + 3*sin(2*pi*1:100*6/100)
x2 = 4*cos(2*pi*1:100*10/100) + 5*sin(2*pi*1:100*10/100)
x3 = 6*cos(2*pi*1:100*40/100) + 7*sin(2*pi*1:100*40/100)
x.3 = x1 + x2 + x3 + rnorm(n = 100, mean = 0, sd = 5)

par(mfrow=c(2,1))
plot.ts(x.1, ylim=c(-20,20), main="Original Sum")
plot.ts(x.3, ylim=c(-20,20), main="Sum with White Noise Term")

x.4 <- x.1 + rnorm(n = 100, mean = 0, sd = 5)
P = Mod(2*fft(x.3)/n)^2
Fr = 0:99/100
plot(Fr, P, type="o", xlab="frequency", ylab="periodogram", main = "Periodogram with White Noise Term")
```


5) The scaled periodogram below is based on re-centered (mean = 0) monthly passengers from Philadelphia International Airport. The most important frequency seems to be around .08, with other spikes occuring around 0.02, 0.25, and 0.41. The large spike at a frequency of about .08 suggests a regularly occuring pattern of about 1/12 = .08333, which makes sense since the plot is based on monthly data.



```{r, echo = FALSE, fig.height=12, fig.width=12}
pax2 <- phl$pax - mean(phl$pax)
n <- length(pax2)
I <- abs(fft(pax2))^2/104
p <- (4/104)*I[1:(104/2)]
f <- 0:51/104
period <- plot(f,p,type='l',xlab='Frequency',ylab='Scaled Periodogram of Monthly Passengers')

```
