---
title: "Brendan Graham HW 6 MAT8444"
author: ""
date: "February 24, 2016"
output: html_document
---

```{r, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}

library(ggplot2)
library(astsa)

#2
#read in data and plot
wins <- read.table("http://www19.homepage.villanova.edu/jesse.frey/Math8444/Phillies2015.txt", 
                   header=FALSE, sep="", na.strings="NA", dec=".", strip.white=TRUE)

names(wins)[1] <- "year"
names(wins)[2] <- "win"

(ts.plot <- ggplot(wins, aes(x=year, y=win)) + geom_point() + geom_line() + 
  labs(x="\nYear", y="Win Percentage\n", title = "Phillies Win Percentage: 1901-2015'n", size = 14))


#format as time series
wins2 <- wins
wins2[1] <- NULL
wins2 <- ts(wins2, frequency = 1)

#plot acf and pacf
plot(wins2, type="o")
diff <- diff(wins2)
plot(diff, type="o")
acf2(wins2)
acf2(diff)

uplim=4
aicmat=matrix(double((uplim+1)^2),uplim+1,uplim+1)
for (i in 0:uplim){
  for (j in 0:uplim){
    aicmat[i+1,j+1]=arima(wins2,order=c(i,0,j))$aic}}
aicmat

```


3. A time series plot of the Phillies win percentage from 1901-2015 is shown in the plot below, along with ACF and PACF plots. The time series plot doesn't indicate any heterogeneity of variance over time or apparent trend, and the ACF plot doesn't indicate much  seasonality, so no differencing seems necessary.

```{r, echo = FALSE, warning=FALSE, message=FALSE}

plot(wins2, type="o")

```


```{r, echo = FALSE, message=FALSE, warning=FALSE, results='asis'}

print(acf2(wins2, main = "ACF and PACF Plots"))

```

The ACF plot indicates a 4 term moving average would be good to use since after a lag of 4 the time series appears like white noise, and the PACF plot indicates 1 AR term may be good to use for the same reason. Fitting this model gives the following results:

```{r, echo=TRUE, warning=FALSE}

model.1 <- sarima(wins2, 1, 0, 4, details=F)

model.1
```

The standardized residuals and the ACF of the residuals look like white noise, the normal Q-Q plot indicates the residuals are normally distributed and the p-values are all very high. The aic value is -288.45. To see if the number of terms can be reduced while still maintaining a low aic value, I created a matrix using the code from the class website which indicates I could lower the aic value by including fewer terms.

```{r, echo = FALSE}

aicmat

```

It seems like a model with 2 auto-regressive terms and no moving average terms would produce the highest aic value.

```{r} {r, echo=TRUE, warning=FALSE}

model.2 <- sarima(wins2, 2, 0, 0, details=F)

model.2
```

This is my preferred model since it has lower aic value and all of the diagnostic plots look the same or better. The only diagnostic plot that could be improved is the normal Q-Q plot, which currently shows some skewness, especially in the lower tail.

The full equation is X~t~ = 0.5295*X*~t-1~ + 0.1768*X*~t-2~ + *W*~t~.

Using the model to predict the Phillies win percentage in 2016 indicates the win percentage will be 0.422 in 2016.

```{r, echo = FALSE, warning=FALSE}

sarima.for(wins2, 1, 0, 0, 0, 2, 0, 0, 1)

```



